\chapter{Introduction}


 \fei{What is the problem here?  Why is it hard?}
 \fei{IMportant: use a picture and a motivating example to demonstrate and show the above points in a realistic setting.  This will help reader understand the problem, how it occurs in real life and why it's hard.}
\fei{Briefly describe 2-3 related work techniques (high level) and state why they are not sufficient, state technical weaknesses.  I suggest you focus on NormA, SAND, DAMP.}


\section{Problem Domain and Motivation}

A time series is an ordered sequence of observations collected over time. Time-series data appear in many real-world applications, including electrocardiogram (ECG) monitoring, electricity demand, and financial activity \citep{schmidl2022anomaly}. In these scenarios, the temporal order of the data is essential. A value is not interpreted only by its magnitude, but also by when it occurs and how it relates to earlier and later observations.

Time-series anomaly detection aims to identify observations or subsequences that do not belong to expected behavior \citep{chandola2009anomaly}. In practice, anomalies may appear in different forms. A point anomaly may be an isolated spike in a sensor reading. A subsequence anomaly may be a short interval whose shape is unusual, even if each individual point is not extreme. A contextual anomaly may be abnormal only under a particular condition, such as a high heart rate during rest but not during exercise. These examples show that anomaly detection is not simply a matter of finding large values; it requires understanding temporal context.

The problem becomes complex when the time series contains concept drift. Concept drift refers to a change in the underlying data distribution over time \citep{gama2014survey,lu2019learning}. In time-series anomaly detection, this means that the definition of normal behavior is not fixed. For example, electricity usage may change between weekdays and weekends, weather measurements may shift gradually across seasons, and ECG patterns may change when a patient's activity or physiological condition changes. A changed pattern may therefore be either a true anomaly or a legitimate new normal state.

Figure~\ref{fig11} illustrates this ambiguity with a simplified example. The first region represents an old normal pattern. Later, the signal gradually changes into a new normal pattern. A detector that compares all future points against only the old normal pattern may incorrectly flag the drift and new normal region as anomalous. However, a detector that adapts without control may also miss the true anomaly. This is the main challenge addressed in this report: anomalies and drift can look similar in the signal, but they require different interpretations.

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
    width=0.98\textwidth,
    height=6.0cm,
    xlabel={Time},
    ylabel={Value},
    xmin=0, xmax=102,
    ymin=-1.7, ymax=2.7,
    axis lines=left,
    xtick=\empty,
    ytick=\empty,
    clip=false,
    label style={font=\large},
    tick label style={font=\small}
]

\addplot[draw=none, fill=blue!7] coordinates {(0,-1.7) (35,-1.7) (35,2.6) (0,2.6)} \closedcycle;
\addplot[draw=none, fill=orange!10] coordinates {(35,-1.7) (60,-1.7) (60,2.6) (35,2.6)} \closedcycle;
\addplot[draw=none, fill=green!8] coordinates {(60,-1.7) (100,-1.7) (100,2.6) (60,2.6)} \closedcycle;

\addplot[blue, thick, domain=0:35, samples=160]
{0.35*sin(deg(0.55*x)) + 0.06*sin(deg(2.7*x))};

\addplot[orange!90!black, thick, domain=35:60, samples=220]
{0.30*sin(deg(0.55*x))*(1-(x-35)/25) + 0.55*sin(deg(1.1*x))*((x-35)/25) + 0.028*(x-35)};

\addplot[green!55!black, thick, domain=60:100, samples=220]
{0.55*sin(deg(1.1*x)) + 0.75 + 0.06*sin(deg(3*x))};
\node[blue, font=\large] at (axis cs:17.5,-1.35) {old normal};
\node[orange!80!black, font=\large] at (axis cs:47.5,-1.35) {drift};
\node[green!45!black, font=\large] at (axis cs:80,-1.35) {new normal};

\end{axis}
\end{tikzpicture}
\caption{Illustration of anomaly detection under concept drift.}
\label{fig11}
\end{figure}

Existing time-series anomaly detection methods provide important foundations for this problem. A common workflow in these methods is to first assign an anomaly score to each point, and then convert these scores into binary anomaly labels using a threshold. NormA learns a normal model from representative subsequences and scores new subsequences according to their distance from this model \citep{boniol2021norma}. SAND extends subsequence anomaly detection to streaming data and updates its model as new observations arrive \citep{boniol2021sand}. DAMP is designed for scalable streaming anomaly detection and uses Matrix Profile ideas to identify left-discords efficiently \citep{lu2022damp}. The final detection result still depends strongly on threshold selection.

Choosing the threshold is difficult in practice. A fixed global threshold may work for one dataset or one normal pattern, but it may fail when the data contain drift. When the normal behavior changes, the anomaly-score distribution may also change. As a result, a strict threshold may incorrectly classify drift-driven normal behavior as anomalous, while a loose threshold may miss true abnormal anomalies. This issue affects baseline methods such as NormA, SAND, and DAMP, because their anomaly scores must still be interpreted through a decision threshold.

AnDri addresses this problem by treating anomaly detection and concept drift detection as related tasks \citep{park2025andri,park2025adaptive}. Instead of assuming that one fixed normal model and one manually chosen threshold are sufficient for the whole time series, AnDri supports an interactive feedback process for threshold adjustment. The user can inspect suspicious candidates, provide labels or corrections, and refine the threshold used to convert anomaly scores into final predictions. Importantly, this feedback-based thresholding workflow can also be applied to baseline methods in the system. For example, even when the anomaly scores are produced by NormA, the user can still participate in the review process and adjust the threshold before final labels are generated. This makes the system useful not only as an AnDri demonstration, but also as an interactive environment for comparing and calibrating different anomaly detection methods.

\section{Project Goals}

The goal of this project is to build and explain an interactive system for anomaly and drift co-detection in time-series data. The system should not only compute anomaly scores, but also help the user understand why a region is suspicious and whether it is more likely to be a true anomaly or a drift-driven normal change.

The first goal is to support co-exploration of anomalies and drift. Co-exploration means that the user can inspect suspicious regions together with their surrounding time-series context, normal-pattern information, anomaly scores, thresholds, and baseline results. This is useful because the same visual change may have different meanings depending on whether it is temporary, recurring, or part of a sustained transition.

The second goal is to provide a visual workflow for long time series. The system allows the user to upload or select a dataset, choose training regions, mark anomaly examples, review uncertain candidates, and compare final results. This workflow is important because many time-series datasets contain tens or hundreds of thousands of points, making manual inspection difficult without interactive zooming, selection, and summary views.

The third goal is to enable adaptive threshold adjustment based on user feedback. Most anomaly detection methods, including AnDri and baseline methods such as NormA, SAND, and DAMP, produce anomaly scores before applying a threshold to decide which points or subsequences are anomalous. However, this threshold is difficult to set manually because score distributions vary across datasets, methods, and normal patterns. The system therefore allows the user to review candidate anomalies, provide feedback, and adjust the threshold dynamically. In AnDri, this process supports pattern-aware interpretation under concept drift. For baseline methods, the same feedback process allows users to calibrate fixed-threshold results more carefully instead of accepting a single automatic cutoff.

\section{Contributions}

The normal-pattern learning component used by AnDri, including Adjacent Hierarchical Clustering, belongs to the broader AnDri research work and is treated in this report as prior methodology. It is not my individual contribution. My individual contributions focus on the surrounding system implementation, interaction workflow, visualization, evaluation support, and deployment.

\begin{enumerate}
    \item \textbf{Interactive AnDri demonstration system.} I implemented and integrated a browser-based system that supports the end-to-end workflow for time-series anomaly and drift exploration. The system connects data loading, training-region selection, anomaly labeling, candidate review, algorithm execution, and result comparison in one interface.

    \item \textbf{Visual analytics workflow for anomaly and drift co-exploration.} I designed the front-end workflow that allows users to inspect long time series, mark anomalous regions, review uncertain candidates, and compare detection behavior across methods. This interaction design makes the detection process interpretable.

    \item \textbf{Backend integration and feedback-based threshold support.} I integrated the Flask backend with AnDri scoring and baseline methods, including NormA, SAND, and DAMP where available. I also implemented the data flow needed for user-label fitting, candidate selection, threshold adjustment, and final prediction generation. This feedback process supports AnDri results as well as baseline outputs, allowing users to calibrate detection thresholds interactively rather than relying only on a fixed global cutoff.

    \item \textbf{Experimental presentation and baseline comparison.} I prepared the system outputs used to compare AnDri with baseline methods using precision, recall, F1-score, and AUC. The experiments focus on ECG time-series data with labeled anomalies and drift regions, showing how AnDri helps distinguish true abnormal behavior from changing normal patterns.

    \item \textbf{Deployment experience.} I deployed the project on a school server using a production-style web serving pipeline. This work helped make the demonstration accessible through a browser and gained practical experience with server configuration, runtime dependencies, static files, and application deployment.
\end{enumerate}

\section{Report Organization}

This report is organized as follows. Chapter 2 introduces the background concepts needed to understand the project, including time-series data, anomaly types, concept drift, subsequence similarity, normal patterns, and active learning. Chapter 3 reviews related work, with a focus on NormA, SAND, DAMP, concept drift adaptation, and benchmark resources. Chapter 4 describes the methodology used in the AnDri workflow, including anomaly-score generation, adaptive thresholding, active learning, and pattern-specific thresholds. Chapter 5 presents the system design and implementation, including the front-end interface, Flask backend, data pipeline, result integration, and deployment environment. Chapter 6 reports the experimental setup, evaluation metrics, ECG results, and analysis. Chapter 7 concludes the report and discusses future work.

