\chapter{Related Work}

This chapter reviews prior work related to time-series anomaly detection and concept drift adaptation. Existing subsequence anomaly detection methods provide strong foundations for detecting unusual time-series regions. After applying the algorithms introduced in this chapter, each point in the time series data is assigned an anomaly score. The higher the score, the more likely it is to be an anomaly. This section briefly explains the key ideas, advantages and disadvantages of the NormA, SAND, DAMP and Andri algorithms.
\fei{Again, anytime you reference and introduce for the first time these existing methods you have to cite them because reader does not know them.}
\reply{Add citation here}{This section briefly explains the key ideas, advantages and disadvantages of the NormA\cite{boniol2021norma}, SAND\cite{boniol2021sand}, DAMP\cite{lu2022damp} and Andri\cite{park2025adaptive} algorithms.}

\section{NormA}

NormA is an unsupervised method for subsequence anomaly detection in time series \citep{boniol2021norma}. The key idea is that normal behavior tends to appear repeatedly over time. NormA first extracts subsequences from the time series and uses offline clustering to identify representative groups of recurring subsequences. The representative subsequences from these clusters form a normal model that characterizes the dominant normal behavior in the data.

During the detection process, the candidate subsequences are compared with the normal patterns that have been learned in the model. When a subsequence differs significantly from the representative normal pattern, it receives a higher anomaly score. In this way, NormA regards the anomaly as a subsequence that deviates from the repetitive patterns discovered from the data.

NormA learns the regular patterns that can summarize repetitive shapes. Therefore, the abnormal score is calculated based on normal behavior. This makes NormA very useful when the data contains repetitive normal patterns, as this method can capture the main structure of the time series and use it as a stable reference standard for subsequent anomaly detection.

However, NormA builds its normal model offline and treats the learned patterns as a relatively fixed description of normal behavior. This assumption may not hold in the presence of concept drift, where normal patterns evolve over time. Patterns that are normal only during a specific period may be underrepresented in the learned model, while outdated patterns can keep affecting detection results. Consequently, changes caused by concept drift may be incorrectly identified as anomalies. To address this limitation, AnDri introduces a dynamic normal model that adapts to changing patterns and updates the set of active normal behaviors over time \citep{park2025adaptive}.


\section{SAND}

SAND is an unsupervised online method for streaming subsequence anomaly detection \citep{boniol2021sand}. SAND is designed for data streams, where new values arrive continuously and the model must be updated over time. SAND maintains a model of normal behavior and scores new subsequences according to their distance from this model.

SAND first builds an initial normal model from an initial batch of the time series. It extracts subsequences from the batch and applies a time-series clustering method based on \(k\)-Shape to group similar subsequences. The centroids of these clusters represent common subsequence shapes and are stored as the current model of normal behavior. When a new batch arrives, SAND clusters the new subsequences and compares the new cluster centroids with the existing model. If a new cluster matches an existing normal pattern, the model can be updated. If it is far from the existing model, the corresponding subsequences receive higher anomaly scores.

SAND supports streaming anomaly detection. Since the model is updated batch by batch, SAND can adapt to changes in the data distribution. This makes it suitable for real-time monitoring tasks. However, SAND mainly adapts the normal model through streaming cluster updates. It does not explicitly represent normal patterns as active, inactive, or newly added concepts.

\section{DAMP}

DAMP is an online anomaly detection method based on the Matrix Profile and discord discovery \citep{lu2022damp}. DAMP detects abnormal situations by identifying subsequences that are significantly different from those previously observed. For each incoming subsequence, DAMP searches for its closest and non-overlapping match in the past. If the most recent previous subsequence is still far apart, the current subsequence is considered abnormal. This distance is used as the abnormal score. In the terminology of matrix contours, DAMP focuses on the left inconsistent items, that is, a subsequence is considered abnormal because it has no approximate matching items in the earlier subsequences of the flow. A larger left nearest neighbor distance indicates that the subsequence is likely abnormal.

DAMP is designed for large-scale and rapidly arriving time series streams, and its performance in anomaly detection far exceeds that of traditional difference search methods. This makes DAMP a useful benchmark for evaluating whether a system can efficiently detect anomalous subsequences in an online environment.

However, DAMP does not explicitly model normal behavior as a set of evolving patterns. When a new normal pattern appears because of concept drift, it may initially look like a discord because it has no close match in the past. As a result, DAMP may assign high anomaly scores to drift-driven changes before the new behavior becomes repeated enough to look normal.

\section{AnDri}

AnDri is an anomaly detection algorithm designed for time series with concept drift \citep{park2025andri,park2025adaptive}. Unlike the methods that use fixed conventional models, Andri assumes that normal behavior changes over time. Therefore, it maintains a dynamic conventional model, in which the regular patterns can be in an active state, an inactive state, or newly added as the time sequence changes.

This algorithm first learns normal patterns from the initial training region. It uses the adjacent hierarchical clustering (AHC) method to group similar subsets while simultaneously considering their temporal proximity. This allows the algorithm to identify normal patterns occurring within a given period, rather than assuming that all normal patterns must appear frequently throughout the entire time series. During the detection process, each new subseries is compared to the currently active normal patterns. If the subsequence is similar to the active pattern, it is regarded as part of the current normal behavior. If an old pattern becomes relevant again, it can be reactivated. If recent subsequences do not match any of the existing active patterns but persist within the recent window, Andri may consider them as new normal patterns caused by concept drift. Otherwise, these subsequences are regarded as abnormal candidate patterns. The anomaly score is calculated based on the distance between the current subsequence and the most similar active normal pattern. This is different from methods such as NormA, which compare subsequences with a more static set of learned normal patterns. By focusing on the active normal behaviors, Andri can more accurately distinguish between true abnormal behaviors and changes in normal behaviors caused by drift.

The main advantage of Andri is to explicitly model the lifecycle of normal patterns. This makes it more suitable for time series data where normal behaviors change, disappear, or reappear over time. In this report, Andri is regarded as the algorithm foundation of the AnomDrift,  \fei{Review all your naming and remove Platform throughout the report.} while subsequent chapters focus on the implemented interactive workflow, visualization, threshold adjustment, and system deployment.
\reply{Review and Change 'AnomDrift Platform' to 'AnomDrift'}{Change 'AnomDrift Platform' to 'AnomDrift'}

\section{Anomaly Score Generation}
\label{sec:ASG}

Anomaly detection algorithms mentioned in this report generate anomaly scores for each data point
\[S=[s_1,s_2,\ldots,s_n],\]
where a larger score indicates stronger evidence of abnormality. In the implementation, scores are normalized to the interval \([0,1]\).

\textcolor{blue}{For NormA, the score is generated by comparing a candidate subsequence with a set of representative normal subsequences learned from the data \citep{boniol2021norma}. Let \[\mathcal{N}=\{N_1,N_2,\ldots,N_K\}\] be the set of normal patterns learned by NormA. For a candidate subsequence \(T_{i,\ell}\), the anomaly score can be written as
\[s_i=\min_{N_k\in\mathcal{N}} d(T_{i,\ell},N_k),\]
where \textit{d} is a distance function between subsequences. If the candidate subsequence is close to one of the learned normal patterns, the score is low. If it is far from all learned normal patterns, the score is high. It is observed that NormA scores anomalies according to their distance from a set of normal subsequence patterns.}

\textcolor{blue}{For SAND, the score is also based on the distance between a candidate subsequence and representative normal shapes, but the normal model is maintained in a streaming setting \citep{boniol2021sand}. Let \(\mathcal{C}_t=\{C_1,C_2,\ldots,C_M\}\) denote the normal cluster centroids or representative shapes maintained by SAND at time \(t\). The anomaly score can be written as
\[s_i=\min_{C_k\in\mathcal{C}_t} d(T_{i,\ell},C_k).\]
A low score means that the candidate subsequence is close to the current streaming normal models. A high score means that it does not match the currently maintained normal patterns.}

\textcolor{blue}{For DAMP, the score is generated from the distance between the current subsequence and its nearest non-overlapping match in the past \citep{lu2022damp}. Let \(\mathcal{H}_i\) be the set of historical subsequences that occur before \(T_{i,\ell}\). The score can be written as
\[s_i=\min_{T_{j,\ell}\in\mathcal{H}_i} d(T_{i,\ell},T_{j,\ell}), \quad j+\ell<i.\]
If the current subsequence has a similar previous match, the score is low. If no similar historical subsequence exists, the nearest-neighbor distance is large, and the subsequence receives a high anomaly score. }

\textcolor{blue}{For AnDri, the score is generated by comparing the current subsequence with the active normal patterns instead of all historical patterns \citep{park2025adaptive}. Let
\[\mathcal{A}_t=\{A_1,A_2,\ldots,A_m\}\]
be the set of active normal patterns at time \(t\). The anomaly score can be written as
\[s_i=\min_{A_k\in\mathcal{A}_t} d(T_{i,\ell},A_k).\]
This means that the score is determined by the closest active normal pattern. If the current subsequence is similar to one active pattern, the score is low. If it is far from all active normal patterns, the score is high.}

\textcolor{blue}{
AnDri also assigns each point to the normal pattern that best matches it:
\[g(i)=\arg\min_{k} d(T_{i,\ell},A_k),\]
where \(g(i)\in\{1,2,\ldots,m\}\) is the index of the closest active normal pattern. The value \(g(i)\) records which normal-pattern group point \(i\) belongs to. This is important for AnomDrift because later threshold refinement can be applied separately to different normal-pattern groups. Baseline methods provide the score sequence \(S\), while AnDri provides both the score sequence \(S\) and the normal-pattern assignment \(g(i)\). This additional information supports pattern-specific thresholding under concept drift.}




AnDri algorithm also assigns each point to a normal pattern index:
\[g(i) \in \{1,2,\ldots,m\},\]
where \(g(i)\) indicates the normal pattern that best matches point \(i\). 

\fei{What is this section trying to say?  What is $S$, and what are the $s_i$?  How are the scores generated for each method?  You have to describe these technical details!  What is the goal of the normal pattern index $g$?  This section has to be strengthened with much more technical deatils of how anomaly scoring works for each method, and their differences.}

\reply{Explain each algorithm in detail}{Please see the blue texts above}