\chapter{Methodology}

    \fei{You cannot use his work, but have to write in your own words.  If you are going to talk about AnDri which is not your work, then describe this in Background chapter.  Move 3.2 to Background. }
    \fei{Describe your methodology using a picture to show what is the problem, and then a picture showing your solution (framework).  This chapter is all about the algorithm details so show using a picture and explain how your algorithm works.  I don't  understand how 3.5 is an algorithm method.}
    \fei{Why is this interesting as an algorithm?  }

This chapter describes the methodological workflow used in this project to detect anomalies in time series with concept drift. The chapter builds on the prior AnDri method, which provides the dynamic normal-pattern model and the anomaly scoring mechanism \citep{park2025andri,park2025adaptive}. The Adjacent Hierarchical Clustering component is treated as part of the prior AnDri method and is not claimed as an individual contribution of this report. The focus here is the end-to-end detection methodology implemented in the demo system: score generation, normal-pattern-aware thresholding, active-learning-based label refinement, and final anomaly decision making.


\section{Problem Definition}

Let \(T=[x_1,x_2,\ldots,x_n]\) be a uni-variate time series, where \(x_i\) is the observed value at time index \(i\). The goal of anomaly detection is to decide whether each time point is normal or anomaly. Instead of making this decision directly from the raw value \(x_i\), an anomaly detection method first assigns each time point an anomaly score \(s_i\). A larger score means that the point is more suspicious.

The detector therefore transforms the original time series into a point-level anomaly-score sequence:
\[S=[s_1,s_2,\ldots,s_n].\]
Here, every point \(i\) has its own anomaly score \(s_i\), and these scores are arranged in the same temporal order as the original observations. In the Andri system interface, this score sequence is visualized as an anomaly-score curve below the original time-series plot.

Although the final score is represented at the point level, many time-series anomaly detection methods compute this score using local temporal context. Given a window length \(\ell\), a local subsequence starting at index \(i\) can be written as
\[T_{i,\ell}=[x_i,x_{i+1},\ldots,x_{i+\ell-1}].\]
Local subsequences are compared with learned normal patterns and then assigned scores to the corresponding points.

After the anomaly scores are generated, the final binary anomaly label is obtained by applying a threshold:
\[
\hat{y}_i =
\begin{cases}
1, & \text{if } s_i > \theta,\\
0, & \text{otherwise,}
\end{cases}
\]
where \(\hat{y}_i=1\) means that point \(i\) is predicted as anomaly, and \(\hat{y}_i=0\) means that it is predicted as normal point. This workflow is common in time-series anomaly detection. Time series anomaly detect methods produce anomaly scores. Final anomaly labels will depend on how the threshold is selected.

Choosing this threshold is difficult when concept drift is present. A fixed global threshold may work in one part of the time series but fail in another part after the normal behavior changes. If the threshold is too low, drift-driven normal behavior may be incorrectly marked as anomalous. If the threshold is too high, true anomalies may be missed. Therefore, the problem is not only how to compute anomaly scores, but also how to interpret and calibrate these scores under changing normal patterns.

AnDri addresses this problem by supporting interactive threshold adjustment. The user can review candidate anomalous points, provide feedback. So that the system can refine the threshold according to the user feedback to convert anomaly scores into final labels. This feedback process is not limited to AnDri's own scoring method. In the demo system, baseline methods such as NormA, SAND, and DAMP can also generate point-level anomaly scores, and the user can still participate in the feedback process to adjust the threshold before final predictions are produced.

Figure~\ref{fig41} illustrates this problem formulation. The upper plot shows the original time series. The lower plot shows the anomaly score of each time point. The threshold is applied to the anomaly-score curve.

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
    name=rawseries,
    width=0.95\textwidth,
    height=4.2cm,
    xmin=0, xmax=100,
    ymin=-1.6, ymax=2.4,
    axis lines=left,
    xtick=\empty,
    ytick=\empty,
    ylabel={Value},
    title={Original time series},
    label style={font=\large},
    title style={font=\large}
]


\def\series{0.45*sin(deg(0.35*x))+ 0.20*sin(deg(1.4*x))+ 1.2*exp(-0.15*(x-50)^2)}

\addplot[
    blue,
    thick,
    domain=0:46,
    samples=150
]
{\series};


\addplot[
    red,
    thick,
    domain=46:54,
    samples=120
]
{\series};

\addplot[
    blue,
    thick,
    domain=54:100,
    samples=150
]
{\series};

\node[
    red!80!black,
    font=\small
] at (axis cs:56,1.85)
{Anomalies};

\end{axis}

\begin{axis}[
    name=scoreseries,
    at={(rawseries.south west)},
    yshift=-4.8cm,
    anchor=north west,
    width=0.95\textwidth,
    height=4.2cm,
    xmin=0, xmax=100,
    ymin=0, ymax=1.8,
    axis lines=left,
    xlabel={Index},
    ylabel={Anomaly score},
    xtick=\empty,
    ytick=\empty,
    title={Anomaly-score curve},
    label style={font=\large},
    title style={font=\large}
]
\addplot[black, thick, domain=0:100, samples=260]
{0.25 + 0.08*sin(deg(0.25*x)) + 1.05*exp(-0.08*(x-50)^2)};

\addplot[dashed, red!70!black, thick] coordinates {(0,0.85) (100,0.85)};

\node[red!70!black, font=\small, anchor=south west] at (axis cs:2,0.88) {threshold \(\theta\)};
\node[red!80!black, font=\small] at (axis cs:50,1.45) {\(s_i > \theta\)};
\end{axis}

\end{tikzpicture}
\caption{ Each point in the original time series is assigned an anomaly score, and the threshold is applied to the anomaly-score curve to obtain final anomaly labels.}
\label{fig41}
\end{figure}




\section{Method Overview}

The implemented workflow has four main stages. First, the system converts the input time series into anomaly scores. For AnDri, this stage also returns the normal pattern associated with each point. Second, the user selects one or more training regions and marks a small number of anomalies. Third, the system estimates anomaly thresholds from the score distribution of the user-labeled points. Fourth, when the score distribution is ambiguous, the system requests additional user feedback for candidate points and updates the threshold iteratively.


\section{Anomaly Score Generation}

Each detection method produces a score sequence
\[S=[s_1,s_2,\ldots,s_n],\]
where a larger score indicates stronger evidence of abnormality. In the implementation, scores are normalized to the interval \([0,1]\).

For AnDri, each point is also assigned to a normal pattern index:
\[g(i) \in \{1,2,\ldots,m\},\]
where \(g(i)\) identifies the normal pattern that best explains point \(i\). This mapping is important because different normal patterns may have different score distributions. A score that is high for one normal pattern may be ordinary for another. AnDri sets thresholds separately for the normal-pattern groups observed in the training-sets instead of using one single global threshold.

For comparison, NormA, SAND, and DAMP are also integrated into the same workflow \citep{boniol2021norma,boniol2021sand,lu2022damp}. These methods generate normalized anomaly scores. The baseline algorithms applies one adaptive threshold, while AnDri applies normal-pattern-specific thresholds.

\section{Adaptive Threshold with User Labels}

The system maintains a label flag \(f_i\) for each point. Initially, all the points are unlabeled. A user can mark points as anomaly or normal points:
\[
f_i =
\begin{cases}
0, & \text{unlabeled},\\
1, & \text{user-marked anomaly},\\
2, & \text{user-confirmed normal}.
\end{cases}
\]
The purpose of the adaptive thresholding step is to estimate a threshold \(\theta\) from the user-marked anomaly scores. The final decision is
\[
\hat{y}_i =
\begin{cases}
1, & s_i > \theta,\\
0, & s_i \leq \theta.
\end{cases}
\]

The demo also allows the user to provide an expected anomaly proportion \(\rho\). This value is used to compute a score cutoff \(q_{\rho}\), corresponding to the boundary of the top \(\rho\) percent of scores in the selected training region. This prevents the threshold from becoming too low when only a small number of anomaly labels are provided.

To model the marked anomaly scores, the backend fits a two-component Bayesian Gaussian mixture model:
\[p(s)=\sum_{k=1}^{2}\pi_k \mathcal{N}(s;\mu_k,\sigma_k^2).\]
Only mixture components with weight greater than a small threshold are treated as active. If fewer than two active components remain, the labeled anomaly scores are considered stable, and the threshold is set using both the mixture statistics and the anomaly-proportion cutoff:
\[\theta = \max(q_{\rho}, \mu-\sigma).\]
Here \(\mu\) and \(\sigma\) describe the active anomaly-score component. This rule keeps the threshold consistent with the expected anomaly proportion while also adapting to the user-confirmed anomalies.




\section{Active Learning for Ambiguous Scores}

When two active Gaussian components are found, the labeled scores are ambiguous. This means the current user labels do not yet separate normal and abnormal score regions clearly. The system then computes the intersection between adjacent Gaussian components:
\[\pi_a \mathcal{N}(x;\mu_a,\sigma_a^2) = \pi_b \mathcal{N}(x;\mu_b,\sigma_b^2).\]
Scores near this intersection are the most uncertain because the two components assign them similar likelihood. The system searches for unlabeled points in this uncertainty region and asks the user to classify them as anomalies or normal points. To keep the interaction manageable, the implementation limits each review round to at most 20 candidate points. If too many points are found, the uncertainty window is narrowed; if too few are found, it is widened.

The active-learning loop has two correction steps. First, if a point previously marked as an anomaly has a low score below the top-\(\rho\) cutoff, the system asks whether it should be removed from the anomaly set. This catches possible false anomaly labels. Second, the system asks the user to label points near the mixture intersection, because these points have the greatest effect on the threshold. After each feedback round, the mixture model is refit and the threshold is updated. The loop stops when the score distribution becomes stable or when no additional uncertain candidates are found.

\section{Pattern-Specific Thresholds for AnDri}

The AnDri workflow extends the thresholding procedure by estimating a separate threshold for each normal pattern represented in the training set. Let \(A\) be the set of user-selected training indices. For each normal pattern \(k\), the training indices assigned to that pattern are
\[A_k = \{i \in A \mid g(i)=k\}.\]
The system fits the same Bayesian mixture and active-learning procedure on the scores in \(A_k\), using only the user labels associated with that normal-pattern group. This produces a threshold \(\theta_k\) for normal pattern \(k\).

The final AnDri decision is therefore
\[
\hat{y}_i =
\begin{cases}
1, & s_i > \theta_{g(i)},\\
0, & s_i \leq \theta_{g(i)}.
\end{cases}
\]
This design is important under concept drift. If the time series contains several normal patterns, each normal pattern may have a different score range. A single global threshold may over-flag a drifting normal region or under-detect anomalies in a lower-variance region. Pattern-specific thresholds allow the decision boundary to follow the changing normal context learned by AnDri.

If a point belongs to a normal pattern that was not present in the user-selected training region, the implementation gets the threshold from the most similar trained normal pattern. The nearest trained pattern is selected by Euclidean distance between normal-pattern subsequences:
\[k^*=\arg\min_{j \in K_{\text{train}}} \|NM_k-NM_j\|_2,\]
and the missing threshold is assigned as \(\theta_k=\theta_{k^*}\). This fallback allows the system to still make a decision for regions not directly covered by user feedback.

\section{Overall Workflow}

The overall methodology can be summarized as follows.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    font=\scriptsize,
    node distance=0.6cm and 1.15cm,
    process/.style={
        rectangle,
        rounded corners,
        draw,
        align=center,
        text width=4.1cm,
        minimum height=0.85cm
    },
    smallprocess/.style={
        rectangle,
        rounded corners,
        draw,
        align=center,
        text width=2.65cm,
        minimum height=0.85cm
    },
    decision/.style={
        diamond,
        draw,
        aspect=2.1,
        align=center,
        inner sep=1pt,
        text width=2.55cm
    },
    arrow/.style={-{Latex}, thick }
]

\node[process] (input)
{Input time series \(T\), training regions, partial user labels, and anomaly proportion \(\rho\)};

\node[process, below=of input] (score)
{Compute normalized anomaly scores \(S=[s_1,\ldots,s_n]\)};

\node[decision, below=of score] (method)
{Using AnDri?};

\node[smallprocess, right=of method] (andri)
{Use normal-pattern assignment\\
\(g(i)\) for each point};

\node[smallprocess, below=of andri] (andrigroup)
{Form pattern-specific groups 
\(A_k=\{i\in A \mid g(i)=k\}\)};

\node[process, below=of method] (baselinegroup)
{Form one global threshold group};

\node[process, below=1.0cm of baselinegroup] (fit)
{Fit or refit Bayesian Gaussian mixture to current user-marked anomaly scores};

\node[decision, below=of fit] (stable)
{Stable score distribution?};

\node[smallprocess, right=of stable] (query)
{Select review candidates:\\
low-score marked anomalies or\\
points near Gaussian intersection};

\node[smallprocess, below=of query] (feedback)
{User confirms candidates as anomaly or normal};

\node[smallprocess, below=of feedback] (update)
{Update labels};

\node[process, below=of stable] (threshold)
{Set threshold for this group\\
\(\theta=\max(q_{\rho},\mu-\sigma)\)};

\node[decision, below=of threshold] (more)
{More groups\\to process?};

\node[process, below=of more] (label)
{Assign final labels using each point's group threshold\\
\(\hat{y}_i=1\) if \(s_i>\theta_{c(i)}\)};

\draw[arrow] (input) -- (score);

\draw[arrow] (score) -- (method);

\draw[arrow] (method) -- node[above] {yes} (andri);

\draw[arrow] (andri) -- (andrigroup);

\draw[arrow] (andrigroup.south) |- ([yshift=10pt]fit.east);

\draw[arrow] (method) -- node[left] {no} (baselinegroup);

\draw[arrow] (baselinegroup) -- (fit);

\draw[arrow] (fit) -- (stable);

\draw[arrow] (stable) -- node[left] {yes} (threshold);
\draw[arrow] (stable) -- node[above] {no} (query);

\draw[arrow] (query) -- (feedback);

\draw[arrow] (feedback) -- (update);
\draw[arrow] (update.east) -- ++(1.2,0) |- ([yshift=-8pt]fit.east);
\draw[arrow] (threshold) -- (more);
\draw[arrow] (more) -- node[left] {no} (label);

\draw[arrow] (more.west) -- ++(-1.15,0) node[left] {yes} |- (fit.west);

\end{tikzpicture}
\caption{Workflow Overview.}
\label{fig42}

\end{figure}


This methodology connects the algorithmic score output with the interactive system workflow. The backend produces normalized scores and, for AnDri, normal-pattern assignments. The frontend uses those values to guide user review, update thresholds, and present final anomaly decisions. In this way, the system does not rely only on a static cutoff. Instead, it uses partial human feedback to adapt the decision boundary and uses AnDri's normal-pattern structure to make the threshold sensitive to concept drift.