\chapter{Conclusion}

This report introduces the AnDri system, which is an interactive web-based platform used for jointly detecting anomalies and drifts in univariable time series data. The report first presents the basic concepts required for the project, including time series anomalies, concept drift, normal patterns, subsequence similarity, and active learning. It then reviews relevant anomaly detection methods, such as NormA, SAND, and DAMP, and discusses the limitations of these methods when the normal behavior evolves over time. Based on this, the methodology section explains how to convert point-level anomaly scores into final labels through threshold selection, and how to optimize this process using user feedback. The main contribution of this project lies in the development and integration of the AnDri system. This system supports data set upload, training area selection, anomaly labeling, detector execution, candidate result review, threshold optimization, final result checking, and comparison with baseline methods. The front end is implemented using JavaScript and Apache ECharts for interactive visualization, while the back end uses Flask to handle detector execution, score normalization, label fitting, similar pattern search, and JSON communication with the browser. Additionally, the system is deployed on the school's server, using a deployment mode based on NGINX and Gunicorn, enabling the demonstration functionality to surpass the local development environment for access and exercising the author's software engineering skills. The experiments show that user-confirmed labels can improve the conversion effect from anomaly scores to final binary labels, although this experiment should be regarded as an idealized feedback environment. Case studies of RCG and weather datasets further show that AnDri helps visually distinguish true anomaly regions from normal behavior changes caused by drift. Compared to baseline methods, the system provides interpretable results because users can view anomaly scores, thresholds, normal patterns, candidate regions, and comparison results in the same workflow. Overall, this project provides a practical workflow that enables users to review uncertain cases, adjust thresholds, compare different methods, and understand why a certain region is determined as abnormal or normal in the context of drift. This makes the system suitable for time series analysis scenarios.



\section{Future Work}
% \fei{At least 3 future work items with a few sentences describing initial ideas of how to solve and why it's important. }

Three directions for future work are provided here. 

First, AnDri can be extended from uni-variate time series to multivariate time series. Many real-world applications involve multiple related variables that evolve together. For example, financial monitoring may include stock prices, volume, and interest rates. Extending AnDri to multivariate data would allow the system to capture relationships among variables and detect anomalies that are only visible when multiple signals are analyzed together.

Second, the adaptive-threshold workflow can be improved with a clearer stopping criterion. In the current system, the threshold is refined through user feedback on candidate anomalies and candidate normal points. This human-in-the-loop design improves interpretability and can reduce false positives, but it may also require repeated user interaction. Future work can define automatic stopping rules based on  confidence in the candidate labels or the expected benefit of asking for another user label. This would help balance detection accuracy with labeling effort.

Third, AnDri should be evaluated on more real-world drift scenarios. The drift regions in ECG dataset are injected artificially. Synthetic drift is helpful for evaluation because the ground truth is known, but it may not fully represent the complexity of real world environments. Future experiments should include more naturally occurring drift in real world scenarios. This would make the evaluation more convincing.